FM project

### 2023-05-07 draft1.Rmd

### Introduction

@@

### Data preparation and overview

@@

```{r, Reliance on Russian oil_tsibble example, echo=FALSE, results='markup',warning=FALSE,message=FALSE}

library(tidyverse)
library(readr)
library(lubridate)
library(tsibble)

oecd_relrus %>% slice_sample(n=10)

In case of other three data sets

```{r,Crude oil petroleum products and nat gas tsibble examples ,echo=FALSE, results='markup',warning=FALSE,message=FALSE}
library(tidyverse)
library(readr)
library(lubridate)
library(tsibble)

@@

oecd_crude_oil <- ts_crude_oil %>%
  filter(`Destination country` %in% oecd_europe) %>% 
  filter(Date >= as.Date("2013-01-15"), Date <= as.Date("2023-01-15"))

@@

oecd_oil_prod <- ts_oil_prod %>% 
  filter(`Destination country` %in% oecd_europe) %>% 
  filter(Date >= as.Date("2013-01-15"), Date <= as.Date("2023-01-15"))

oecd_oil_prod %>% slice_sample(n=10)

@@

oecd_nat_gas <- ts_nat_gas %>%
  filter(`Destination country` %in% oecd_europe) %>% 
  filter(Date >= as.Date("2013-01-15"), Date <= as.Date("2023-01-15"))

oecd_nat_gas %>% slice_sample(n=10)
```

**Placeholder for methodologies**

### 2. Time series graphics

Done

@@

```{r, decomposition, echo=FALSE, results='markup',warning=FALSE,message=FALSE}

```{r, decomposition, echo=FALSE, results='markup',warning=FALSE,message=FALSE}

@@

@@

```
#### 3.3. Decomposition

```{r, decomposition, echo=FALSE, results='markup',warning=FALSE,message=FALSE}

dcmp_oecd_nat_gas <- oecd_nat_gas_agg %>%
  model(stl = STL(`Amount of Natural gas (MMcf)`))

components(dcmp_oecd_nat_gas)
```

```{r}

>>>>

### 1.Introduction

@@

### 2. Time Series Graphics

In this section of the report, we will be creating time series graphics using various techniques such as time plots, seasonal plots, seasonal subseries plots, and ACF(). These plots will provide insights into the different time series in the dataset and help us understand their characteristics, such as seasonality, trend, and cyclicity. We will also look for unusual observations and changing patterns over time and explore the differences between the series in terms of these features. Additionally, for this analysis, we have decided to focus on three specific time series: US exports of Natural Gas to OECD, US Exports of Crude Oil to OECD, and US exports of Petroleum products to OECD. We will create the same types of graphs for all of these time series, including time plots, seasonal plots, seasonal subseries plots, and ACF plots. These time series span the last ten years from 2013 to 2023. Through these analyses, we aim to gain insights into the seasonality, cyclicity, trend, unusual observations, changing patterns, and differences between these time series, and determine if any of the series resemble a white noise process. By analyzing these time series graphics, we can gain a better understanding of the data and identify any potential patterns or anomalies that may exist within it.

### 2.1 Dataset

@@

In this section, we will analyze three datasets that pertain to US exports of Crude oil, Petroleum Products, and Natural Gas, all measured in barrels. Prior to converting the data to tsibble format, we performed data cleaning and formatting. To simplify the data, we converted each dataset to have the same structure, with three columns: date, destination country, and the amount exported. After standardizing the format, we converted each dataset to tsibble format. We then filtered the data to only include countries that are part of the OECD and summarized the results based on the amount exported on each date, removing country names and leaving only the date and amount exported. Finally, we merged all three tsibble tables into one by using the left join function and pivoting the data, resulting in a final time series that represents US exports of the three different products in barrels to OECD countries. The tsibble contains three columns: Date, Export Type, and Amount.

```{r, data_prep, echo=FALSE, results='markup',warning=FALSE,message=FALSE}
library(fpp3)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(tidyr)
library(lubridate)

@@

oecd_crude_oil <- ts_crude_oil %>%
  filter(`Destination country` %in% oecd_europe) %>% 
  filter(Date >= as.Date("2010-01-15"), Date <= as.Date("2023-01-15"))

@@

oecd_oil_prod <- ts_oil_prod %>% 
  filter(`Destination country` %in% oecd_europe) %>% 
  filter(Date >= as.Date("2010-01-15"), Date <= as.Date("2023-01-15"))

################################################################################
#Reliance on Russian gas

@@

oecd_nat_gas <- ts_nat_gas %>%
  filter(`Destination country` %in% oecd_europe) %>% 
  filter(Date >= as.Date("2010-01-15"), Date <= as.Date("2023-01-15"))

oecd_nat_gas_agg <- oecd_nat_gas_agg %>% 
  mutate(`Amount of Natural gas (BOE)` = `Amount of Natural gas (MMcf)` * 1000 / 6000 )

oecd_nat_gas_agg <- oecd_nat_gas_agg %>% 
  select(Date, `Amount of Natural gas (BOE)`)

oecd_merged <- left_join(oecd_nat_gas_agg, oecd_oil_prod_agg, by = "Date") %>%
  left_join(oecd_crude_oil_agg, by = "Date")

total_exp_oecd <- oecd_merged %>%
  pivot_longer(cols = -Date, names_to = "Export Type", values_to = "Amount")

total_exp_oecd %>% ggplot(
 aes(Date,Amount,color = `Export Type`)) +
   geom_line()+
   labs(x = "Date", y = "Amount")+
  theme(legend.position = "bottom")

@@

```{r, box_cox, echo=TRUE, results='markup',warning=FALSE,message=FALSE}

@@

#### 3.3. Decomposition

```{r, decomposition, echo=TRUE, results='markup',warning=FALSE,message=FALSE }

dcmp_oecd_crude_oil <- oecd_crude_oil_agg %>%
  model(stl = STL(`Amount of Crude Oil (Thousand Barrels)`))

components(dcmp_oecd_crude_oil)

dcmp_oecd_oil_prod <- oecd_oil_prod_agg %>%
  model(stl = STL(`Amount of total Petroleum Porducts (Thousand Barrels)`))

components(dcmp_oecd_oil_prod)

dcmp_oecd_nat_gas <- oecd_nat_gas_agg %>%
  model(stl = STL(`Amount of Natural gas (BOE)`))

components(dcmp_oecd_nat_gas)

@@

```{r, decomposition_plots, echo=FALSE, results='markup',warning=FALSE,message=FALSE }

components(dcmp_oecd_crude_oil) %>% autoplot()

components(dcmp_oecd_oil_prod) %>% autoplot()

components(dcmp_oecd_nat_gas) %>% autoplot()
```
